Interpretable Neural Networks using EAGGA

### 2025-03-09 paper/main.tex
little bit of restructuring

regularisation
- LM: ridge (L2) + lasso (L1)
  * ridge: overcome multi-collinearity by pulling coeffs close to 0
  * lasso: variable selection by setting coeffs = 0
- NN:
  * L2 regularisation implicitly included in SGD when using weight decay (pull params towards 0) -> does this combat multi-collinearity as well or serve another purpose?
  * L1
    + key claim of DL: feature selection obsolete, network finds important ones itself
    + still, feature usually always has some impact (back this up), no explicit feature selection
    + explicit feature selection simply by not inluding a feature
  * another classical NN regularisation technique: dropout against co-adaptation
  * early stopping against overfitting when using iterative method

@@

\subsection{Hyperparameter Optimization (HPO)}
- grid + random search
- Bayesian optimisation
-> both unsuitable for our task
  + would be ok for weight decay (L2) + dropout optim
  + but: approaches for feature selection (L1) not feasible
    (a) 0-1 penalty on included features added to loss, problems:
      - how to choose regularisation parameter?
      - not all features equally important -> weighted 0-1 loss? how to choose weights?
    (b) directly optimise over binary hpyercube -> how to optim over non-ordinal discrete space?
-> evolutionary algorithms can be a solution for these types of problems
  * CMA-ES
  * NSGA-II
    - non-dominated sorting
    - crowding distance
    -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals
\subsection{Neural Architecture Search (NAS)}
not sure if super relevant as I don't use a specific NAS technology (e.g. one-shot, etc.) but simply optimise over \# layers + \# nodes as HPs

\subsection{Multi-Objective Optimization}
- problem: all aforementioned approaches are a-priori approaches -> we choose best HPs (weight decay, features used, dropout) w.r.t. performance
- intro MOO
\subsubsection{Scalarisation}
\subsubsection{Pareto-optimality}

%likely shouldn't be an entire section
\section{EAGGA}
- NSGA-II inspired evolutionary / genetic algorithm
- introduces group structure for more efficient optimization over the entire search space

@@

\section{Conclusion and Future Outlook}

Here is a citation \cite{EAGGA}.

@@

\appendix
\section{Software used}
for implementation we used openml \cite{OpenML}, \cite{OpenMLPython}, numpy \cite{numpy}, pandas \cite{pandas1}, \cite{pandas2}, pytorch \textbf{TODO},
scikit-learn \cite{scikit-learn}, scipy \cite{SciPy}, pymoo \cite{pymoo}, and tqdm \cite{tqdm}

>>>>

regularisation
- LM: ridge (L2) + lasso (L1)
  * ridge: overcome multi-collinearity by pulling coeffs close to 0
  * lasso: variable selection by setting coeffs = 0
- NN:
  * L2 regularisation implicitly included in SGD when using weight decay (pull params towards 0) -> does this combat multi-collinearity as well or serve another purpose?
  * L1
    + key claim of DL: feature selection obsolete, network finds important ones itself
    + still, feature usually always has some impact (back this up), no explicit feature selection
    + explicit feature selection simply by not inluding a feature
  * another classical NN regularisation technique: dropout against co-adaptation
  * early stopping against overfitting when using iterative method
-> all of those require specifying Hyperparameters in advance

@@

\subsection{Hyperparameter Optimization (HPO)}
- grid + random search
- Bayesian optimisation
- evolutionary algortihms
  * CMA-ES
  * differential evolution

\subsection{Neural Architecture Search (NAS)}
give background on NAS (e.g. one-shot, etc.), but for this project wanted to have a "minimal viable product" to test whether NNs can yield competitive performance
on tabular ML + be interpretable using EAGGA at all before optimising the inner networks to the max

\subsection{Multi-Objective Optimization}
- problem: all aforementioned approaches are a-priori approaches -> we choose best HPs (weight decay, features used, dropout) w.r.t. performance
- intro MOO
\subsubsection{Pareto-optimality}
\subsubsection{A-priori}
\subsubsection{A-posteriori}
- NSGA-II (MOEA)
  * non-dominated sorting
  * crowding distance
  -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals

%likely shouldn't be an entire section
\section{EAGGA}
- for our problem (interpretability) we cannot use regular EA algorithms as we are also interested in reducing pairwise interactions + non-monotone feature effects
- this creates high-dim search space as shown in EAGGA paper -> they provide a framework to efficiently traverse through a space that is reduced to only sensible HP
  configuration (e.g. interaction of A,B with A monot incr and B monot decr -> couldn't be guaranteed -> don't look at it in first place)
- NSGA-II inspired evolutionary / genetic algorithm
- introduces group structure for more efficient optimization over the entire search space

@@

\section{Conclusion and Future Outlook}
- MO BO on group structure space possible?
- here is a citation \cite{EAGGA}.

@@

\appendix
\section{Software used}
for implementation we used openml \cite{OpenML}, \cite{OpenMLPython}, numpy \cite{numpy}, pandas \cite{pandas1}, \cite{pandas2}, pytorch \cite{PyTorch},
scikit-learn \cite{scikit-learn}, scipy \cite{SciPy}, pymoo \cite{pymoo}, and tqdm \cite{tqdm}
